---
title: "Build and deploy a stroke prediction model using R"
date: "`r Sys.Date()`"
output: html_document
author: "Put your name!"
---

# About Data Analysis Report

This RMarkdown file contains the report of the data analysis done for the project on building and deploying a stroke prediction model in R. It contains analysis such as data exploration, summary statistics and building the prediction models. The final report was completed on `r date()`. 

**Data Description:**

According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

This data set is used to predict whether a patient is likely to get stroke based on the input parameters like gender, age, various diseases, and smoking status. Each row in the data provides relevant information about the patient.


# Task One: Import data and data preprocessing

## Load data and install packages

```{r}
# Install and load required libraries
if (!require("tidyverse")) install.packages("tidyverse", dependencies = TRUE)
if (!require("caret")) install.packages("caret", dependencies = TRUE)
if (!require("data.table")) install.packages("data.table", dependencies = TRUE)
if (!require("GGally")) install.packages("GGally", dependencies = TRUE)

library(tidyverse)
library(caret)
library(data.table)
library(GGally)
```


## Describe and explore the data

```{r}
# Load dataset
data <- fread("healthcare-dataset-stroke-data.csv")

# View structure and summary
str(data)
summary(data)

# Handle missing values
data <- data %>%
  mutate_if(is.numeric, ~ifelse(is.na(.), median(., na.rm = TRUE), .))

# Convert the outcome variable to a factor
data$stroke <- as.factor(data$stroke)

# Check the levels to ensure proper conversion
levels(data$stroke) <- c("No", "Yes")

# Convert categorical variables to factors
data <- data %>%
  mutate(across(where(is.character), as.factor))

# Scale numeric columns
setDT(data)
numeric_cols <- names(data)[sapply(data, is.numeric)]
data[, (numeric_cols) := lapply(.SD, scale), .SDcols = numeric_cols]

set.seed(123)
trainIndex <- createDataPartition(data$stroke, p = 0.8, list = FALSE)
train_data <- data[trainIndex, ]
test_data <- data[-trainIndex, ]

```



# Task Two: Build prediction models

```{r}
set.seed(123)


logistic_model <- train(
  stroke ~ ., 
  data = train_data, 
  method = "glm", 
  family = "binomial", 
  trControl = trainControl(method = "cv", number = 5)
)
summary(logistic_model)

```

```{r}

```



# Task Three: Evaluate and select prediction models

```{r}
set.seed(123)
dt_model <- train(
  stroke ~ ., 
  data = train_data, 
  method = "rpart", 
  trControl = trainControl(method = "cv", number = 5)
)
print(dt_model$finalModel)
rpart.plot::rpart.plot(dt_model$finalModel)

```

```{r}
set.seed(123)
svm_model <- train(
  stroke ~ ., 
  data = train_data, 
  method = "svmRadial", 
  preProcess = c("center", "scale"),
  trControl = trainControl(method = "cv", number = 5)
)
summary(svm_model)

```

```{r}
# Predictions and confusion matrices
logistic_preds <- predict(logistic_model, test_data)
dt_preds <- predict(dt_model, test_data)
svm_preds <- predict(svm_model, test_data)

# Evaluate performance
logistic_cm <- confusionMatrix(logistic_preds, test_data$stroke)
dt_cm <- confusionMatrix(dt_preds, test_data$stroke)
svm_cm <- confusionMatrix(svm_preds, test_data$stroke)

# Summarize metrics
evaluation <- data.frame(
  Model = c("Logistic Regression", "Decision Tree", "SVM"),
  Accuracy = c(logistic_cm$overall["Accuracy"], dt_cm$overall["Accuracy"], svm_cm$overall["Accuracy"]),
  Sensitivity = c(logistic_cm$byClass["Sensitivity"], dt_cm$byClass["Sensitivity"], svm_cm$byClass["Sensitivity"]),
  Specificity = c(logistic_cm$byClass["Specificity"], dt_cm$byClass["Specificity"], svm_cm$byClass["Specificity"])
)

print(evaluation)

if (!require("pROC")) install.packages("pROC", dependencies = TRUE)
library(pROC)

logistic_roc <- roc(test_data$stroke, as.numeric(predict(logistic_model, test_data, type = "prob")[,2]))
svm_roc <- roc(test_data$stroke, as.numeric(predict(svm_model, test_data, type = "prob")[,2]))

plot(logistic_roc, col = "red", main = "ROC Curves")
plot(svm_roc, col = "blue", add = TRUE)
legend("bottomright", legend = c("Logistic Regression", "SVM"), col = c("red", "blue"), lwd = 2)

```


# Task Four: Deploy the prediction model

```{r}
# Save the chosen model
saveRDS(logistic_model, "stroke_prediction_model.rds")

# Load the model
loaded_model <- readRDS("stroke_prediction_model.rds")

```




# Task Five: Findings and Conclusions
































